---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(GGally)
library(broom)
```


```{r}
juice <- read_csv("data/orange_juice.csv") %>% clean_names()

view(juice)
summary(juice)
```

# There sems to be redundant variables storeId, store7 and STORE all provide the same information
```{r}
#removing store & store7
juice <- juice %>%
  select(-c(store7, store))
         
```

```{r}
#create new column for whether MM was purchased
juice <- juice %>% 
  mutate(purchase_MM = ifelse(purchase == "MM", T, F )) %>% 
  select(-purchase)


juice <- juice %>% 
  mutate(special_ch = ifelse(special_ch == "1", T, F )) %>% 
  mutate(special_mm = ifelse(special_mm == "1", T, F )) #%>% 

```


```{r}
#check for other potential redundant variables

alias(glm(purchase_MM ~ ., data = juice, family = binomial(link = 'logit')))

# sale_price_mm, sale_price_ch, price_diff, list_price_diff cqn be computed from other variables and will be removed
# also rmoved pct_disc_mm & pct_disc_ch

juice <- juice %>%
  select(-c(sale_price_mm, sale_price_ch, price_diff, list_price_diff, pct_disc_ch, pct_disc_mm))

juice <-  juice %>% 
  mutate(store_id = as.factor(store_id))

view(juice)

```

```{r}
#plot variables against purchase MM

juice %>% 
  ggplot(aes(x = purchase_MM, y = weekof_purchase)) +
    geom_boxplot()
```

```{r}

juice %>% 
ggplot() +
  geom_bar(aes(x = purchase_MM, fill = store_id), position = "dodge")
```

#It seems as thought there is a link between the store and the purchase of MM


```{r}
#the price of Citrus Hill
juice %>% 
  ggplot(aes(x = purchase_MM, y = price_ch)) +
    geom_boxplot()
```
# There seems to be little difference made by the price of Citrus Hill and whether Minute Maid is purchased


```{r}
#the price of Minute Maid
juice %>% 
  ggplot(aes(x = purchase_MM, y = price_mm)) +
    geom_boxplot()
```

#As expected the price of Minute Maid determines whether it is purchased


```{r}
#the discount on CH
juice %>% 
  ggplot(aes(x = purchase_MM, y = disc_ch)) +
    geom_boxplot()
```
#As expected has an inverse relationship with purchase of MM


```{r}
#the discount on MM
juice %>% 
  ggplot(aes(x = purchase_MM, y = disc_mm)) +
    geom_boxplot()
```
#As expected impacts whether MM is purchased


```{r}
#special on CH
juice %>% 
ggplot() +
  geom_bar(aes(x = purchase_MM, fill = special_ch))
```
#As expected this has an inverse relationship


```{r}
juice %>% 
ggplot() +
  geom_bar(aes(x = purchase_MM, fill = special_mm))
```

#As expected a special on MM increases MM purchases

```{r}
#Loyalty to CH
juice %>% 
  ggplot(aes(x = purchase_MM, y = loyal_ch)) +
    geom_boxplot()
```
Clearly the greater the loyalty to CH, the less likely to buy MM


```{r}
#decided to exclude weekofpurchase variable for simplicity
juice_model <- glm(purchase_MM ~ store_id +  price_mm + disc_ch + disc_mm + special_ch + special_mm + loyal_ch,  data = juice, 
                   family = binomial(link = 'logit'))

```


```{r}
tidy(juice_model)
```

Based on the output above store_id & special_ch & special_mm are not significant

