---
title: "R Notebook"
output: html_notebook
---

```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)
library(GGally)
library(janitor)


```


```{r}
titanic_set <- read_csv("data/titanic_decision_tree_data.csv")

shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```



1.1 Question 1


```{r}
#Take only observations which have a survived flag (i.e. that aren’t missing)
titanic_set <- titanic_set %>% 
  filter(!is.na(survived))

#Turn your important variables into factors (sex, survived, pclass, embarkation)

titanic_set <- titanic_set %>% 
mutate_if(is.character, as.factor)


#Create an age_status variable which groups individuals under (and including) 16 years of age into a category called “child” category and those over 16 into a category called “adult”.

titanic_set <- titanic_set %>% 
  mutate(age_status = if_else(age <= 16, "child", "adult" )) %>% 
  filter(!is.na(age_status))



#Drop any variables you don’t need (X1, passenger_id, name, ticket, far, cabin)
titanic_set <-  titanic_set %>% 
select(-c(X1, passenger_id, name, ticket, fare, cabin))


#also set age_status to factor
titanic_set <-  titanic_set %>% 
  mutate(age_status = as_factor(age_status))
view(titanic_set)

titanic_set <- titanic_set %>% 
  mutate(survived = if_else(survived == 1, "lived", "died" ))

view(titanic_set)

```


2. Have a look at your data and create some plots to ensure you know what you’re working with before you begin. Write a summary of what you have found in your plots. Which variables do you think might be useful to predict whether or not people are going to die? Knowing this before you start is the best way to have a sanity check that your model is doing a good job.

```{r}

ggpairs(titanic_set)
#this doesn't really show much so choose o plot individually as not too many variables

# check plot for age
titanic_set %>% 
ggplot(aes(x = survived, y = age)) +
  geom_boxplot()
  
#There doesn't seem to be a hugh variation in the ages of those that survived and those that perished
```

```{r}
# check plot for sex
titanic_set %>% 
ggplot(aes(x = survived, fill = sex)) +
  geom_bar()

# clearly a higher proportion of women survived than men - women and children first!!
```

```{r}
# check plot for age
titanic_set %>% 
ggplot(aes(x = survived, fill = age_status)) +
  geom_bar()

#more children survived than died this would be expected as stated above
```

```{r}
titanic_set %>% 
ggplot(aes(x = pclass, fill = survived), position = "dodge") +
  geom_bar()

#class is clearly a decider in whether an individual survived
```

```{r}
titanic_set %>% 
ggplot(aes(x = embarked, fill = survived), position = "dodge") +
  geom_bar()

#where an individual boarded seems to have an impact on whether they survived however from wikipeadia:
#920 passengers boarded Titanic at Southampton – 179 First Class, 247 Second Class, and 494 Third Class
#perhaps mainly first class passengers boarded at those other two ports??
```

```{r}
titanic_set %>% 
ggplot(aes(x = sib_sp, fill = survived), position = "dodge") +
  geom_bar()

# The majority of children had no siblings, I'm not sure this variable has any bearing on whether someone survived
```

```{r}
titanic_set %>% 
ggplot(aes(x = parch, fill = survived), position = "dodge") +
  geom_bar()
## The majority of children had no parents onboard, I'm not sure this variable has any bearing on whether someone survived
```


3. Now you can start to build your model. Create your testing and training set using an appropriate split. Check you have balanced sets. Write down why you chose the split you did and produce output tables to show whether or not it is balanced. [Extra - if you want to force balanced testing and training sets, have a look at the stratified() function in package splitstackshape (you can specify multiple variables to stratify on by passing a vector of variable names to the group argument, and get back testing and training sets with argument bothSets = TRUE)]


```{r}
# get hthe number of rows
(n_data <- nrow(titanic_set))

# create a test sample index thast is 20% of the total sample (142)
test_index <- sample(1:n_data, size = n_data*0.2)

# create test set of 142 rows
titantic_set_test  <- slice(titanic_set, test_index)


# create training set of the remaining data (include all but the test_set)
titantic_set_train <- slice(titanic_set, -test_index)
```

```{r}
# check the balance between the test and training sets

titantic_set_test %>%
 tabyl(survived)

titantic_set_train %>%
 janitor::tabyl(survived)
```

The sets are reasonably balanced:
test set has a survived % of 46%
training set has a survived % of 39%




4. Create your decision tree to try and predict survival probability using an appropriate method, and create a decision tree plot.

```{r}
#create decision tree
titanic_set_fit <- rpart(
  formula = survived ~ ., 
  data = titanic_set, 
  method = 'class'
)

#plot tree
rpart.plot(titanic_set_fit, 
           yesno = 2, 
           #fallen.leaves = TRUE, 
           faclen = 6, 
           digits = 4)
```

