---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(GGally)
library(arsenal)
```

```{r}
diamond <- read_csv("diamonds.csv")

diamond <- diamond %>% 
  select(-X1)

```


2.We expect the carat of the diamonds to be strong correlated with the physical dimensions x, y and z. Use ggpairs() to investigate correlations between these four variables.

```{r}
ggpairs(diamond, columns = c(1, 8:10))
```
```{r}
# ANS: All the correlations are all above 0.9, therefore there is a very strong correlation for each of these variables
```



3. So, we do find significant correlations. Let’s drop columns x, y and z from the dataset, in preparation to use only carat going forward.

```{r}
diamond <- diamond %>% 
  select(-c(x:z))
```


4. We are interested in developing a regression model for the price of a diamond in terms of the possible predictor variables in the dataset.
i. Use ggpairs() to investigate correlations between price and the predictors (this may take a while to run, don’t worry, make coffee or something).

```{r}
ggpairs(diamond)
```
ii. Perform further ggplot visualisations of any significant correlations you find.

```{r}
#The only significant correlation is between carat & price
diamond %>% 
  ggplot(aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```


5. Shortly we may try a regression fit using one or more of the categorical predictors cut, clarity and color, so let’s investigate these predictors:
