---
title: "R Notebook"
output: html_notebook
---

Homework Quiz


1.I want to predict how well 6 year-olds are going to do in their final school exams. Using the following variables am I likely under-fitting, fitting well or over-fitting? Postcode, gender, reading level, score in maths test, date of birth, family income.

```{r}
#'*1.I want to predict how well 6 year-olds are going to do in their final school exams. Using the following variables am I likely under-fitting, fitting well or over-fitting? Postcode, gender, reading level, score in maths test, date of birth, family income.*


# ANS This is likely to be an example of over-fitting

```



```{r}
#'*2.If I have two models, one with an AIC score of 34,902 and the other with an AIC score of 33,559 which model should I use?*

# ANS :The model with the lower AIC score should be used
```


```{r}
#'*3.I have two models, the first with: r-squared: 0.44, adjusted r-squared: 0.43. The second with: r-squared: 0.47, adjusted r-squared: 0.41. Which one should I use?*

# ANS : we would choose the value with the highest adjusted R-squared value

```



```{r}
#'*4.I have a model with the following errors: RMSE error on test set: 10.3, RMSE error on training data: 10.4. Do you think this model is over-fitting?*

#

```


```{r}
#'*5.How does k-fold validation work? *

#split the data into k parts.  Set aside one part as the test set.  Train the model using the data for the other k-1 parts.  We test the data on each part, using the other k-1 parts to train the data.  The error is averaged across all the tests.

```


```{r}
#'* 6.What is a validation set? When do you need one?*

# A validation set is used once as a final test of the expected performance of the model

```


```{r}
#'* 7.Describe how backwards selection works. *

regsubsets()

```


8.Describe how best subset selection works.

9.It is estimated on 5% of model projects end up being deployed. What actions can you take to maximise the likelihood of your model being deployed?

10.What metric could you use to confirm that the recent population is similar to the development population?

11.How is the Population Stability Index defined? What does this mean in words?

12.Above what PSI value might we need to start to consider rebuilding or recalibrating the model

13.What are the common errors that can crop up when implementing a model?

14.After performance monitoring, if we find that the discrimination is still satisfactory but the accuracy has deteriorated, what is the recommended action?

15.Why is it important to have a unique model identifier for each model?

16.Why is it important to document the modelling rationale and approach?




```{r}
library(tidyverse)
library(janitor)
library(GGally)
library(ggfortify)
```

```{r}
avocado <- read_csv("data/avocado.csv") %>% clean_names()

view(avocado)
```

```{r}
#check if there are any variables which can be removed (none reported)
alias(lm(average_price ~ ., data = avocado))
```


```{r}
#check for NAs in datset (there are none)
summary(avocado)
```

```{r}
avocado_subset<- avocado %>% 
  select(-c(region, x1, total_volume, total_bags, x4770, small_bags:x_large_bags))

```

```{r}
# meaning of PLU numbers:
# 4046	AVOCADOS	Hass	Small		60 size and smaller
# 4225	AVOCADOS	Hass	Large		48 and 40 size
# 4770	AVOCADOS	Hass	All Sizes

avocado_subset <- avocado_subset %>% 
mutate(date = str_extract(date, "-[0-9]+-")) %>% 
  mutate(date = str_remove_all(date, "-")) %>%
  mutate(winter = if_else(date == "01" | date == "02" |
                            date == "12", T, F )) %>% 
  mutate(spring = if_else(date == "03" | date == "04" |
                            date == "05", T, F )) %>% 
  mutate(summer = if_else(date == "06" | date == "07" |
                            date == "08", T, F )) %>% 
  mutate(autumn = if_else(date == "09" | date == "10" |
                            date == "11", T, F )) %>% 
  select(-date) %>%
  mutate(conventional = if_else(type == "conventional", T, F )) %>% 
  mutate(organic = if_else(type == "organic", T, F)) %>% 
  select(-type) %>%
  rename(small = x4046) %>% 
  rename(large = x4225)

```

```{r}
#check which predictors are significant for our response variable
ggpairs(avocado_subset)
#this doesn't give a clear indication of what predictors to choose
```


```{r}

model1 <- lm(average_price ~ spring, data = avocado_subset)
summary(model1)

#we can explain 0.6% of the variance in average price based on spring avocado sales.  The P-value is under 0.01 so is statistically significant. The RSE is $0.40.

```


```{r}
model2 <- lm(average_price ~ winter, data = avocado_subset)
summary(model2)
#we can explain 2% of the variance in average price based on winter avocado sales. The P-value is under 0.01 so is statistically significant. The RSE is $0.40.
```


```{r}
model3 <- lm(average_price ~ summer, data = avocado_subset)
summary(model3)
#we can explain 0.5% of the variance in average price based on winter avocado sales. The P-value is under 0.01 so is statistically significant. The RSE is $0.40.
```

```{r}
model4 <- lm(average_price ~ organic, data = avocado_subset)
summary(model4)
#we can explain 38% of the variance in average price based on winter avocado sales. The P-value is under 0.01 so is statistically significant. The RSE is $0.32.
```

```{r}
model5 <- lm(average_price ~ conventional, data = avocado_subset)
summary(model5)
#we can explain 38% of the variance in average price based on winter avocado sales. The P-value is under 0.01 so is statistically significant. The RSE is $0.32.
```


```{r}
model6 <- lm(average_price ~ small, data = avocado_subset)
summary(model6)

#we can explain 4% of the variance in average price based on winter avocado sales. The P-value is under 0.01 so is statistically significant. The RSE is $0.39.
```

```{r}
model7 <- lm(average_price ~ large, data = avocado_subset)
summary(model7)
#we can explain 3% of the variance in average price based on winter avocado sales. The P-value is under 0.01 so is statistically significant. The RSE is $0.40.
```

```{r}
model8 <- lm(average_price ~ year, data = avocado_subset)
summary(model8)
#we can explain 3% of the variance in average price based on winter avocado sales. The P-value is under 0.01 so is statistically significant. The RSE is $0.40.
```

