---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
```


```{r}
books_data <- read_csv("data/books.csv")
```


```{r}
#Inspecting the data
#view(books_data)
```

I've decided to answer the following question:

1. 5 longest & shortest books
2. author with the most books
3. author those works have the highest ratings
4. oldest 5 books, newest 5 books
5. average rating of all books - book that have a rating above average
6. a publication date in the future



However first we need to check and clean up the data

```{r}
#check for duplicate book IDs (duplicate titles exist but I believe that these are unique books as these are published by different organisations - e.g. "Salem's Lot)
books_data %>% 
  distinct(bookID)

#All IDs are disctinct
```



```{r}
#checking for the number of Nas/NaNs in the data
books_data %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
  
```


```{r}
#Looking for columns which have values of 0 where it doesn't make sense
#clearly average_rating, ratings_count & text_reviews_count could have a value of zero

books_data[books_data$bookID == 0,] 
books_data[books_data$title == 0,]
books_data[books_data$authors == 0,]
books_data[books_data$isbn == 0,]
books_data[books_data$isbn13 == 0,]
books_data[books_data$language_code == 0,]
books_data[books_data$num_pages == 0,]
books_data[books_data$publication_date == 0,]
books_data[books_data$publisher == 0,]

#There is a large number of 0 values in the 'num_pages column

```


```{r}
#convert zeros to NA in num_pages column
books_data_nozeroes <- books_data %>% 
  mutate(num_pages = na_if(num_pages, 0))

```

```{r}
#checking the above command has given the desired outcome
books_data_nozeroes3[books_data_nozeroes3$num_pages == 0,]
```


#The titles of the 5 longest & shortest books
```{r}
#The title of the 5 shortest books
books_data_nozeroes %>% 
 slice_min(num_pages, n = 5)
```

```{r}
#The title of the 5 longest books
books_data_nozeroes %>% 
  slice_max(num_pages, n = 5)
```



2. author with the most books
```{r}
books_author <- group_by(books_data, authors)
arrange((summarise(books_author , author_count = n())), desc(author_count))
```

3. author those works have the highest ratings
```{r}
view(books_author)
#Need to find the median rating for all an author's book and then sort these returning the author's name, didn't mange this one

```

4. oldest 5 books, newest 5 books
```{r}
books_data %>% 
 slice_min(publication_date, n = 5)

```
```{r}
books_data %>% 
 slice_max(publication_date, n = 5)

```

5. average rating of all books - book that have a rating above average
```{r}
#remove 0 values from average_rating
books_data_ratings <- books_data %>% 
  mutate(average_rating = na_if(average_rating, 0))

#add a new column to capture average rating for all books
books_data_ratings_avg <- books_data_ratings %>% 
mutate(overall_avg_rating = median(average_rating, na.rm = TRUE))

#compare a books rating to overall average rating
books_data_ratings_avg %>% 
  filter(average_rating > overall_avg_rating)
#try to arrange this in descending order by average_rating like I did in question 2, however it continually failed???
```



6. A publication date in the future
```{r}
#I attempted to convert publication_date to date type and planned to compare it to current date today(), however I couldn't get it to work! I also looked at the lubridate package, however I couldn't get it to work

books_data %>% 
as.Date(publication_date, "%Y-%m-%d")
```


