---
title: "R Notebook"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
library(tidyverse)
```

```{r}
titanic <- read_csv("data/titanic.csv") 
```


# Add a column to confirm whether an individual was part of the crew
```{r}
titanic <- titanic %>% 
  mutate(crew = ifelse(class %in% c("engineering crew", "deck crew",
                                    "restaurant staff", "victualling crew"), TRUE, FALSE))
```


# create variables to allow probabilites to be calculated 
```{r}
(total_crew <- titanic %>%
  filter(crew == TRUE) %>% 
    nrow())

  (crew_survived <- titanic %>% 
    filter(crew == TRUE & survived == "yes") %>% 
    nrow())

#probability that crew survived
(prob_crew_survived <- crew_survived/total_crew)


(total_guests <- titanic %>%
  filter(crew == FALSE) %>% 
    nrow())

  (guests_survived <- titanic %>% 
    filter(crew == FALSE & survived == "yes") %>% 
    nrow())

#probability that guests survived
(prob_guests_survived <- guests_survived /total_guests)


```
ANS: The probabality that guests survived is 0.38, whereas the probabality the crew
survived was 0.24, therefore it was more likely for guests to have survived.


1. If I roll a six-sided dice once, what is the probability of getting a number less than 5?

```{r}
s_roll_die <- nrow(rolldie(1))

s_roll_die_filtered <- rolldie(1) %>% 
  filter(X1 < 5 ) %>% 
  nrow

(probability <- s_roll_die_filtered/s_roll_die)
```

ANS: As there is 6 potential outcomes the probability is 4/6 (0.667)

-------------------------------------------

2.How many outcomes are there in the sample space of an experiment in which we toss three dice?
```{r}
rolldie(3) %>% 
nrow
```

ANS : there are 216 outcomes in the sample sapce when 3 die are rolled

-------------------------------------------

3. In an experiment where we roll four dice:

i. how many outcomes are in the event total_equals_5_6_7 
(i.e. the total of the four dice equals 5,6 or 7)?
ii. what is the probability of this event?
iii. what assumptions do we apply in answering part ii?

```{r}
(total_5_6_7 <-  rolldie(4) %>% 
  mutate(total = X1 + X2 + X3 + X4) %>% 
  filter(total == 5 | total == 6 | total == 7) %>% 
  nrow())
```

ANS for i:  There are 34 outcomes when 4 die are rolled and the anwer is 5, 6, or 7


```{r}
(four_die_outcome <- rolldie(4) %>% 
  nrow())

(prob_5_6_7 <- total_5_6_7/four_die_outcome)

```

ANS for ii. There is 1296 outcomes when rolling 4 die, therefore the probability if the answer being 5, 6 or 7 is 
34/1296 = 0.026

iii. The assumption we make in ii.is that the die are not weighted and therefore the
probability of rolling 1-6 is 1/6



4. Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

```{r}

perms <- as_tibble( 
permutations(
  n = 5,
  r = 5
  )
)

decreasing_height <- perms %>% 
  filter(V1 == 1 & V2 == 2 & V3 == 3 & V4 == 4 & V5 == 5) %>% 
  nrow

(prob_chairs <- decreasing_height/nrow(perms))

```
ANS: The probability is 1/120 or 0.0083


5. I roll six dice:

i.what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?

```{r}

(perms <- rolldie(6) %>% 
  as_tibble)

number_perms <- nrow(perms)

ordered_outcome <- perms %>% 
  filter(X1 == 6 & X2 == 5 & X3 == 4 & X4 == 3 & X5 == 2 & X6 == 1) %>% 
  nrow

(ordered_outcome/number_perms)

```

ANS for i. The probability is 1/46,656 (0.0000214)


ii. what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?


ANS: if the order is not fixed the number will be 6 factorial (which is 720)
therefore probability is 720/46,656 which equals .015

```{r}
720/number_perms

#test that the numerator should be 720
x <- isin(perms, c(6,5,4,3,2,1), ordered = F)
as.data.frame(table(x))

```

6. If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades? 

event A: spade is drawn
event B: black card is drawn
intersection of  A & B  is : 13/52
probability of B is : 26/52 (1/2)

Based on the conditional prob. formula:
```{r}
(13/52)/(1/2)
```
ANS: the probability is 0.5



7. Eight patients: Alice, Bob, Carmen, Dinesh, Ella, Fergus, Gianna and Henry are waiting to see a doctor, but only five appointment slots are left today.

i.In how many different orders can the doctor see patients today?

```{r}
patients <- c("Alice", "Bob", "Carmen", "Dinesh", "Ella", "Fergus", "Gianna", "Henry")
(patient_perms <- permutations(
  n = 8,
  r = 5,
  v = patients,
  repeats.allowed = F
) %>% as_tibble())
```
ANS: there is 6720 different ways to order this


ii. What is the probability of Gianna and Danesh both being seen today?

```{r}
patient_perms_2 <- isin(patient_perms, c("Dinesh", "Gianna"), ordered = F)
as.data.frame(table(patient_perms_2))
```

