---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(janitor)
library(broom)
library(animation)
library(factoextra)

```

```{r}
customers <- read_csv("data/mall_customers.csv") %>% clean_names()
```


```{r}
#remove non-numeric data, add a lable for each customer
customers <- customers %>% 
  mutate(customer_id = str_c("Customer ", customer_id)) %>% 
  select(-gender) %>% 
  column_to_rownames("customer_id")



#view(customers)

```


1. You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings.

```{r}
# plot the data to try and see if there are clusters
customers %>% 
ggplot(aes(x = annual_income_k, y = spending_score_1_100, col = age)) +
  geom_point()
```
ANS: there seems to be 5 clusters, one in each corner and one in the centre.


```{r}
#check if scaling is required :

summary(customers)
#based on the min and max numbers I don't believe that any scaling is required
```


2. We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.
Perform k-means clustering and chose a value of k.

```{r}
(clustered_cust <- kmeans(customers, 
                            centers = 5, 
                            nstart = 25))

# display cluster in more friendly format
tidy(clustered_cust)


```


Visualise the clustering for your chosen value of k.

```{r}
customers %>% 
  kmeans.ani(centers = 5)
```



Do you think the clustering seems a good fit for this data?
```{r}
#'*ANS: this is not the clustering I expected to see based on the initial scatterplot *
```


#check and see if there is an optimal number of clusters:
```{r}
fviz_nbclust(customers, 
             kmeans, 
             method = "wss", 
             nstart = 25)
#'* based on the output from this there is not an optimal number of clusters *

```


```{r}
fviz_nbclust(customers, 
             kmeans, 
             method = "gap_stat", 
             nstart = 25, 
             )
#The output from this suggests that the optimal number of clusters is 1!
```




Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the ‘Segmentation & clustering intro’ lesson).

```{r}
augment(clustered_cust, customers)
```
```{r}
tidy(clustered_cust)
```

Cluster 1 seesm to be medium income, age and spending score.
Cluster 2 seems to be medium age, low income and low spending score
Cluster 3 seems to be low age, high income and high spending score
Cluster 4 seems to be medium age, high income and low spending score
Cluster 5 seems to be low age, low income and high spending score