---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(e1071)
library(skimr)
```

```{r}
phones <- read_csv("20190928-items.csv") %>% 
  clean_names()
```

2. Find the brand with the highest numbers of phones in the dataset.
```{r}
phones %>%
  group_by(brand) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

The brand with the highest number of phones is Samsung with 397

```{r}
phones_samsung <- phones %>%
  filter(brand == "Samsung") 
```



3. For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

```{r}
phones_samsung %>% 
summarise(skewness = skewness(rating, type = 1))
```


The data for ratings is moderately skewed to the left
As this is moderately left skewed the median is preferred to the mean

```{r}
phones_samsung %>% 
  skim(rating)
```


```{r}
f_r = dunif(x = phones_samsung$rating, min = 1, max = 5 )
x <- phones_samsung$rating

ratings <- tibble(
  l = seq(0, 6, by = 0.01),
  f_r = dunif(x = x, min = 1, max = 5)
                )

 ggplot(aes(x = ratings, y = f_r)) +
  geom_line() +
  ylim(0, 1) + 
  xlab("ratings (minutes)") + ylab("f_r (probability density)")

```


