---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
```

Q1. Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.

```{r}
tweets_data <- read_csv("data/code_clan_tweets.csv") %>% clean_names()

```

```{r}
nrow(tweets_data)
```

```{r}
ncol(tweets_data)
```
```{r}
names(tweets_data)
```

Q2. Find the total number of favourited tweets (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
tweets_data %>%
  filter(is_quote == F) %>% 
summarise(favourite_tweets_count = sum(favorite_count))


```


Q3.Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
tweets_data %>% 
  filter(is_quote == F) %>%
  group_by(source) %>% 
  summarise(mean_retweets = mean(retweet_count))


```


Q4. Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.
```{r}
#check if and where NAs exist
tweets_data %>%
  summarise(across(, ~ sum(is.na(.x))))

tweets_data %>% 
  mutate(media_type = coalesce(media_type, "text")) %>%
  group_by(media_type) %>%
  summarise(favourite_tweets_count = sum(favorite_count)) %>%
  arrange(desc(favourite_tweets_count))

```


Q5. Find the mean number of characters that a CodeClan tweet contains. You can either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if possible to see if they match (but it’s not essential).

```{r}
tweets_data %>% 
  summarise(mean_tweet_length = mean(display_text_width))

test <- tweets_data %>% 
  transmute(display_text_width, num_chars = str_length(text))

view(test)
```


Q6.The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.
```{r}
tweet_info <- read_csv("data/code_clan_info.csv") %>% clean_names()

joined_tweets <- inner_join(tweets_data, tweet_info, by = "tweet_id")

#view(joined_tweets)

```


Q7. From your new joined data, create a new tibble codeclan_hashtags containing only the tweet_id and hashtags in lowercase for analysis. Keep only those tweets with hashtags.
```{r}
codeclan_hashtags <- tibble(
    joined_tweets %>% 
    filter(!is.na(hashtags)) %>% 
    transmute(tweet_id, lower_hash = str_to_lower(hashtags))
  )


codeclan_hashtags2 <- joined_tweets%>% 
  select(tweet_id, hashtags) %>%
  mutate(lowcase_hashtag = str_to_lower(hashtags)) %>%
  select(-hashtags) %>%
  drop_na(lowcase_hashtag)

class(codeclan_hashtags2)

```



EXTENSIONS

Q1.Some tweets have more than one hashtag, denoted by the c( symbols at the start of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags tibble from above to find all the cases in which the hashtag string begins with charactersc(.

```{r}
codeclan_hashtags %>%
  mutate(letter_one_two = str_sub(lower_hash, 1, 2)) %>%
  filter(str_detect(letter_one_two, "c\\("))

#can be done this way also:
codeclan_hashtags %>%
  filter(str_detect(lower_hash, "c\\("))

```



Q2.Use the str_detect() function to find all tweets with text that mentions “Edinburgh”, and count how many such tweets there are.
```{r}
joined_tweets %>%
  filter(str_detect(text, "Edinburgh")) %>%
  summarise(count = n())
```



Q3.Use str_extract_all() with a regular expression to find out which Twitter users CodeClan have been tweeting.
```{r}
joined_tweets %>%
  transmute(users_tweeted = str_extract_all(text, "@[:punct:]*[:alnum:]*")) #%>% 
  #filter(!is.na(users_tweeted))

joined_tweets %>%
  transmute(users_tweeted = str_extract_all(text, "@[a-zA-Z0-9_]+")) %>%
  filter(!is.na(users_tweeted))


```

