---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(CodeClanData)
```


Question 1.

Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.

```{r}
backpack <- CodeClanData::backpack
summary(backpack)

```

create a logical column for known back problems for question 6
```{r}
backpack_updated <- backpack %>% 
  mutate(back_problems_logical = if_else(back_problems == 1, TRUE, FALSE))
```



Question 2.

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

```{r}

backpack %>% 
 group_by(backpack_weight) %>% 
 summarise(counts = n()) %>% 
ggplot() +
  geom_smooth(aes(x = backpack_weight, y = counts), se = FALSE) +
  labs(
  x = "Backpack Weight (lb)",
   y = "Number of Students",
   title = "Weight carried by students"
 ) +
  scale_x_continuous(
    breaks = seq(2,40, by = 2)) +
  scale_y_continuous(
    breaks = seq(1,12, by = 1))
  
  
```

```{r}
backpack %>% 
 group_by(backpack_weight) %>% 
 summarise(counts = n()) %>% 
ggplot() +
  geom_col(aes(x = backpack_weight, y = counts), fill = "blue") +
  labs(
  x = "Backpack Weight (lb)",
   y = "Number of Students",
   title = "Weight carried by students") +
  scale_x_continuous(
    breaks = seq(2,40, by = 2)) +
 scale_y_continuous(
    breaks = seq(1,12, by = 1))
```



Question 3.

Are there differences in backpack weight between sex?

```{r}
ggplot(backpack) +
  geom_histogram(aes(x = backpack_weight, fill = sex), col = "black",   
                 show.legend = FALSE, binwidth = 1) +
    facet_wrap(~ sex) +
    labs(
    x = "Backpack Weight (lb)",
    y = "Gender",
    title = "Backpack Weight by Gender"
    ) +
  scale_x_continuous(
    breaks = seq(0,40, by = 5)) +
 scale_y_continuous(
    breaks = seq(1,10, by = 1))
```
Males carry heavier backpacks, however the average backpack size for both sexes
is 10lb

-----------------------------------------------

Question 4.

What is the relationship between body weight and backpack weight?

```{r}
ggplot(backpack) +
  geom_smooth(aes(x = backpack_weight, y = body_weight), se = FALSE) +
   labs(
    x = "Backpack Weight (lb)",
    y = "Body Weight (lb)",
    title = "Weight carried"
    ) +
    scale_x_continuous(
    breaks = seq(0, 40, by = 5)) +
 scale_y_continuous(
    breaks = seq(100, 200, by = 10))
```

As an individual weight increases they carry a heavier bag, this seems to make sense as larger individuals would expect to be capable of carrying a heavier backpack 

----------------------------------------------------------

Question 5.

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.

```{r}
ggplot(backpack) +
  geom_smooth(aes(x = backpack_weight, y = body_weight), se = FALSE) +
  facet_wrap(~ sex) +
   labs(
    x = "Backpack Weight (lb)",
    y = "Body Weight (lb)",
    title = "Weight Carried by Gender"
    ) +
    scale_x_continuous(
    breaks = seq(0, 40, by = 5)) +
 scale_y_continuous(
    breaks = seq(100, 200, by = 10))
```
Females have a lower bodyweight than males and therefore carry smaller backpacks
---------------------------------------------

Question 6.

What proportion of students report back pain?

```{r}
ggplot(backpack_updated) +
  aes(fill = back_problems_logical) +
  geom_bar(aes(x = back_problems_logical, y = stat(prop), group = 1
               )) +
   labs(
    x = "Back Problems",
    y = "Percentage",
    title = "Percentage of individuals with back problems"
    ) 
```
Approx. 65% of students report back pain

-----------------------------------------------

Question 7.

Is this proportion different between men and women?

```{r}
ggplot(backpack_updated) +
  aes(fill = back_problems_logical) +
  geom_bar(aes(x = back_problems_logical, y = stat(prop), group = 1
               )) +
  facet_wrap(~ sex) +
   labs(
    x = "Back Problems",
    y = "Percentage",
    title = "Percentage of individuals with back problems"
    ) 
```
A higher proportion of women (approx. 40%) have back problems, less than 20% (approx.) of men have back problems.

-------------------------------------------------------

Question 8.

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.





```{r}
#I assume these are all to appear in the same graphs, I wasn't sure how to get 
#the different sets of axis aligned, therefore I couldn't get this to work

ggplot(backpack_updated) +
  geom_smooth(aes(x = backpack_weight, y = body_weight), se = FALSE) +
  geom_bar(aes(x = back_problems_logical, y = stat(prop), group = 1, 
               scales = "free"
               )) +
  facet_grid(~ sex) +
   labs(
    x = "Backpack Weight (lb)",
    y = "Body Weight (lb)",
    title = "Weight Carried by Gender"
    ) +
    scale_x_continuous(
    breaks = seq(0, 40, by = 5)) +
 scale_y_continuous(
    breaks = seq(100, 200, by = 10))


```

