---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
library(broom)
library(ggfortify)
```


```{r}
project <- read_csv("data/project_management.csv")
```



i)Plot the data, taking estimated_length as the independent variable and actual_length as the dependent variable.
```{r}
project %>% 
  ggplot(aes(x =  estimated_length, y = actual_length)) +
  geom_point()

```

ii)Calculate the correlation coefficient of estimated_length and actual_length and interpret the value you obtain.

```{r}
project %>% 
  summarise(corr_coefficient = cor(actual_length, 
                                   estimated_length))
```
The correlation coefficient is 0.8047.  This is a 'very strong' positive correlation


iii)Perform a simple linear regression using actual_length as the dependent variable, and estimated_length as the independent variable. Save the model object to a variable.

```{r}
#use lm() to fit a linear model
(model <- lm(formula = actual_length ~ estimated_length, data = project))
```

iv)Interpret the regression coefficient of estimated_length (i.e. slope, gradient) you obtain from the model. How do you interpret the r2 value reported by the model?
```{r}
#add predictions to check output received from linear model
project <- project %>% 
  add_predictions(model)
```


A 1 unit increase in the estimated length changed the actual length by 1.223

```{r}
(glance_output <- glance(model))
```

The coefficient of determination (R-Squared) shows that 64.5% of the variation in actual length can be explained by the variation in estimated length.


v)Is the relationship statistically significant? Remember, to assess this you need to check the p-value of the regression coefficient (or slope/gradient). But you should first check the regression diagnostic plots to see if the p-value will be reliable (don’t worry about any outlier points you see in the diagnostic plots, we’ll return to them in the extension).

```{r}
(tidy_output <- tidy(model))
```



```{r}
autoplot(model)
```

